\section{Introduction}\label{sec:intro}

This paper reports on strategies and tools employed to
run two iterations of a large-scale functional programming and verification course at the Technical University of Munich (TUM).
While the first iteration (winter semester 2019, 1057 participants)
took place on campus,
the second iteration (winter semester 2020, 1031 participants) was affected by the COVID-19 pandemic and took place in virtual space.
Previous iterations of the course were introduced in~\cite{next_1100};
however, we were facing two novel challenges:

\paragraph{Soaring Enrolments}
The relatively young field of computer science has
become one of the largest study programmes around the globe.
The increase of student enrolments is dramatic
\cite{comp_sci_growth1,comp_sci_growth2}
while employment of new teaching staff often lags behind.
At TUM, the number of new enrolments in computer science more than doubled between 2013 and 2021 from 1110 to 2644 (an increase of 138\%)
while academic staff only increased from 439 to 573 (31\%) \cite{tum_numbers}.

This drastic increase not only requires more physical resources -- like larger lecture halls and more library spaces --
but also academic staff for supervision.
Given the discrepancy in growth between student enrolments and staff employment,
automation of supervision and feedback mechanisms is inevitable.
Automation, however, should not
negatively affect the quality of the teaching.

\paragraph{Online Teaching}
The ongoing COVID-19 pandemic forced a radical
transition from on-site teaching to online classes.
Lecturers had to rethink the way they present material and interact with students,
teaching assistants the way they assist students in tutorial sessions.
Students, on the other hand, suffer from a lack of social interaction and communication, leading to higher
levels of stress, anxiety, loneliness, symptoms of depression, and diminished affective engagement~\cite{students_lockdown1,onlineengagement1}.

In our experience, the disconnect between students and lecturers as well as the lack of on campus interaction between students may also lead to \emph{cramming}:
the practice of showing little participation during the semester
while studying extensively just before the exam.
Cramming tends to result in poor long-term retention and shallow understanding of material.
Indeed, the benefit of spacing learning events apart rather than cramming has been demonstrated in hundreds of experiments \cite{cramming1,cramming2}.

\vspace{0.5\baselineskip}\noindent
Besides these general challenges,
there is a third -- subject-specific --
challenge we were keen to tackle:

\paragraph{Functional Programming is Practical}
Feedback by students and personal experience has shown us that many students
at TUM question the applicability and usefulness
of functional languages beyond academia.
They are disappointed by a lack of industrial insight
and real world -- or at least interactive -- applications.
Indeed, some even perceive functional programming as an obstacle;
after all, they already know how to program imperatively.

Good educators do not just teach but inspire:
we have to bring the benefits of functional languages
closer to our students' hearts
by showing real-world applicability and making functional programming fun and engaging.

\paragraph{Contributions}

In this paper,
we present our answers to these challenges
and provide tools and exercises for other educators.
Our contributions are
\begin{enumerate*}[label=\arabic*)]
  \item a description of technical tools that enable large-scale (functional) programming courses in physical and virtual space,
  \item a toolbox of engagement mechanisms
for programming courses
and an evaluation thereof, and
  \item reusable exercises and tools for other functional programming educators.
\end{enumerate*}
Our resources can be found in the central repository of this paper.\footnote{\url{https://github.com/kappelmann/engaging-large-scale-functional-programming/}}

\paragraph{Outline}

We begin by describing the underlying conditions of our course and its syllabus in \cref{sec:course_structure_conditions}.
The tools and teaching methods employed during lectures are explained in \cref{sec:lectures}.
\cref{sec:practical_part} describes the
mechanisms, tools, and technical setup
we used to create an engaging experience
that is scalable
for the practical part of the course.
% including tutorials, homework assignments, and
% competition systems.
It also includes a framework for I/O
testing in Haskell
that solves a transparency issue prevalent in traditional approaches.
\cref{sec:cyp} introduces
``Check Your Proof'' -- a tool created
by our lab to automatically check simple inductive proofs for Haskell programs.
\cref{sec:exam} describes how we adapted our exams to the COVID-19 situation and large number of participants.
Finally, \cref{sec:related_work} outlines related work
and \cref{sec:conclusion} concludes with a summary and aspects to improve.

